### 2021-07-22 code/aa-flowers-pilot_analysis.Rmd
moved preprocessing to script and joined data files

---
title: "flowers-pilot-analysis"
output:
  html_document: default
  pdf_document: default
---

```{r set-up, include=F}
library(tidyverse)
library(jsonlite)
library(here)
source(here("code/preprocessing/00_preprocessing_functions.R"))
theme_set(theme_bw())

##Data import constants
date_start=lubridate::ymd('2021-07-19')
image_location=here("figs")

```

# Pre-process
```{r process-data, message=FALSE, warning=FALSE}
data_list <- c("technical_pilots", "collabmulti_pilot0", 
               "collabmulti_20", "compet_20")
process_all_data(data_list)
join_all_data()
```

```{r read-processed-date, message=FALSE}
d.players <- read_csv(here("data/processed_data/joined_data/players.csv")) %>% distinct()
d.games <- read_csv(here("data/processed_data/joined_data/games.csv")) %>% distinct()
d.rounds <- read_csv(here("data/processed_data/joined_data/rounds.csv")) %>% distinct()
d.raw_chat <- read_csv(here("data/processed_data/joined_data/raw_chat.csv")) %>% distinct()
```

```{r}
#exclude BLOCKS with <12 completed (all 3 participants responded) rounds
rounds_exclude <- d.rounds %>% group_by(gameId, blockNum) %>% tally() %>% filter(n!=36) %>% select(gameId,blockNum)

d.rounds.final <-  d.rounds %>% anti_join(rounds_exclude)
d.raw_chat.final <- d.raw_chat %>% anti_join(rounds_exclude) %>% write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
```

```{r}
game_completion <- d.round_results %>% 
  group_by(gameId) %>% 
  summarize(num_rounds=max(trialNum+1)) %>% 
  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength)) %>%
  mutate(gameComplete = ifelse(num_rounds == 48, TRUE, FALSE))
#knitr::kable(summary)
message("Full games")
game_completion %>% filter(num_rounds==48)
message("Partial games")
game_completion %>% filter(num_rounds!=48) 
```

```{r}
d.chat.raw <- d.round_results.raw %>%
  filter(createdAt >= date_start) %>%
  anti_join(rounds_exclude) %>%
  #there will always be an alert for each active round, so drop NA
  filter(!is.na(chat))%>%
  select(-c(ends_with('response'), ends_with('time'), ends_with('utility'), context, stageIds)) %>%
  #mutate(chat = ifelse(is.na(chat), '{}', chat)) %>%
  rename(roundID = `_id`) %>%
  mutate(chat = map(chat, .f = ParseJSONColumn)) %>%
  unnest(chat) %>%
  arrange(createdAt, gameId, blockNum, repNum) %>% 
  left_join(d.players %>% select(playerId, name)) %>%
  left_join(d.round_results %>% select(gameId, blockNum, repNum, trialNum, playerId, playerResponse, playerUtility)) %>%
  left_join(d.games %>% select(gameId, condition, chatEnabled)) %>%
  mutate("participantAction" = ifelse(type == "alert", paste(name, "selected", playerResponse, "for", playerUtility, "point(s)"), NA)) %>%
  select(gameId, trialNum, condition, chatEnabled, playerId, name, text, participantAction, everything()) %>%
  write_csv(here(data_location, 'raw_chat.csv'))

#extract the alternative flowers
d.contexts <- read_csv(here(data_location,'rounds.csv'),guess_max=10000) %>% 
  mutate(data.context = map(data.context, 
                            .f = ParseJSONColumn)) %>% 
  select("_id":"data.numPlayers") %>%
  mutate(data.context = map(data.context,
                            function(df) df %>% gather(key, value) %>%
                              mutate(key = gsub('.', '-', key, fixed=TRUE)) %>%
                              separate(key, into = c("label", "key"), sep = "-") %>%
                              spread(key, value))) %>%
  unnest(data.context) %>%
  rename_with(~ gsub("data.", "", .x, fixed = TRUE))
```

Recast games with a language USE condition
If there were fewer than 3 lines of communication, cast langUse = False
```{r}
d.games <- d.games %>% left_join(d.chat.raw %>% group_by(gameId) %>% summarize(numChat = sum(!is.na(text))) %>% 
                                   mutate(langUse = (numChat >5)) %>% select(-numChat)) %>%
  mutate(langUse = ifelse(is.na(langUse), FALSE, langUse)) %>%
#add a toddle for completed games
  left_join(game_completion %>% select(gameId, gameComplete))
```

How often do participants select a flower that is hidden from them?

# Demographics
```{r demographics}
d.exit.survey <- read_csv(here(data_location, 'player-inputs.csv')) %>%
  filter(createdAt >= date_start) %>%
  left_join(d.games, by = c('gameId')) %>%
    rename_with(~ gsub("data.", "", .x, fixed = TRUE))
```
```{r feedback}
d.exit.survey %>% select( language, fair, chatUseful, feedback, time) %>% knitr::kable()
```

# Pretty pictures

Correlation between response time and utility?

```{r}
d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameComplete)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  ggplot(aes(x=time_sec, y = as.numeric(playerUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(langUse)) +
  geom_point(alpha = .25) + 
  labs(x = "log time (in seconds)",
       y = "individual utility",
       title = "Response time in sec and utility", subtitle =  "each participant, each round")

d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  ggplot(aes(x=log(time_sec), y = as.numeric(playerUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(langUse)) +
  geom_point(alpha = .25) + 
  labs(x = "log time (in seconds)",
       y = "individual utility",
       title = "Response time in sec (log) and utility", subtitle =  "each participant, each round")
```

Easiest way to do this, is to derive the game start date with the latest participant ready time and the game finishedAt

```{r}
d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameLength, gameComplete)) %>%
  filter(gameComplete) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         full_condition = paste0(condition, "-", chat)) %>%
  group_by(gameId, condition, chat, langUse, full_condition, gameLength) %>%
  summarize(totalUtility = sum(playerUtility))%>%
  ggplot(aes(x=gameLength, y = as.numeric(totalUtility), color = full_condition)) + 
  facet_grid(rows = vars(condition), cols = vars(langUse), scales = "free_y") +
  geom_point(alpha = .75) + 
  labs(x = "Game length (in minutes)",
       y = "total points by group",
       title = "Length of game in mins and utility", subtitle =  "per game")

d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse,gameLength, gameComplete)) %>%
  filter(gameComplete) %>%
  mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
         chatUse = ifelse(langUse, "chatUse", "noChat"),
         full_condition = paste0(condition, "-", chatUse)) %>%
  group_by(gameId, condition, chat, full_condition,gameLength) %>%
  summarize(totalUtility = sum(playerUtility))%>%
  mutate(totalUtility = ifelse(condition=="coopCartel", totalUtility/3, totalUtility)) %>%
  ggplot(aes(x=as.numeric(gameLength, units="mins"), y = as.numeric(totalUtility), color = full_condition)) + 
  #facet_grid(rows = vars(condition), cols = vars(chat), scales = "free_y") +
  geom_point(alpha = .75) + 
  geom_smooth(method=lm, alpha=.3)+
  labs(x = "Game length (in minutes)",
       y = "total points by group",
       title = "Length of game in mins and utility", subtitle =  "per game")

ggsave(paste0(image_location, "/game_length.png"))
```

```{r chat}
#group raw chat by participant
d.chat <- read_csv(here(data_location, "raw_chat.csv"), guess_max=1000) %>% 
  filter(type == "message") %>%
  full_join(d.round_results, c("gameId", "trialNum", "repNum", "playerId", "numPlayers", "blockNum")) %>%
 # filter(!is.chitchat) %>% 
  mutate(text = gsub("\\n", '', fixed = T, text),
         text = gsub("[/?/.]", ' ', text),
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1) %>% 
  group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
  summarize(text = paste0(text, collapse = ', '),
            total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars)) %>%
  anti_join(rounds_exclude)

```

Pick just one game to look at, make sure the plot is working?
```{r}
condition = "competCartel"

d.chat %>% 
         left_join(d.games %>% select(gameId, createdAt,condition, langUse, chatEnabled, gameComplete)) %>% 
         filter(langUse, gameComplete) %>%
ggplot(aes(x = repNum, y=total_num_words, color = as.factor(playerId))) + 
  geom_jitter(alpha = .5) + 
  facet_grid(cols = vars(blockNum), rows = vars(gameId), scales = "free_y") + 
  geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Cooperative + Lang Used", subtitle = "Number of words", y="Number of words", x="Round number", color="playerId")

ggsave(paste0(image_location, "/words_over_rounds.png"))
```

Accuracy
sum utility for entire group/max possible utility per group (or, the max flower *3)
Each condition is going to have a different maximum utility

coopcartel will have a max utility = max flower * 3 (players) * 3 (incentive)

competcartel will have a max utility = sum(top_3_flowers)
```{r}
d.max_utility <- d.contexts %>% 
  left_join(d.games %>% select(gameId, condition)) %>%
  mutate(utility= as.numeric(utility)) %>%
  group_by(gameId, blockNum, repNum) %>% 
  slice_max(order_by =utility,n = 3)%>%
  summarize(competCartel = sum(utility),
            coopMulti= sum(utility),
            coopCartel = max(utility * 9)) %>% ungroup() %>%
  pivot_longer(cols = c(competCartel, coopMulti, coopCartel), names_to = "condition", values_to = "maxUtility")

d.utility <- d.round_results.final %>%
         left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>% 
  group_by(gameId, blockNum, repNum, trialNum,condition, langUse) %>% 
  summarize(group_utility = sum(as.numeric(playerUtility)))%>% ungroup() %>%
  left_join(d.max_utility) %>%
  mutate(prop_utility = group_utility/maxUtility,
         langUsed = if_else(langUse, "Lang", "Nonlang"))

d.utility %>% 
  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
  geom_point(alpha = .4) + 
  facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")

ggsave(paste0(image_location, "/accuracy.png"))

d.utility %>% 
  ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
  geom_point(alpha = .4) + 
  facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")

ggsave(paste0(image_location, "/accuracy.png"))

d.utility %>% 
  filter(langUsed == "Lang") %>%
  ggplot(aes(x = repNum, y=prop_utility, color = as.factor(gameId))) + 
  geom_point(alpha = .4) + 
  facet_grid(cols = vars(blockNum), rows = vars(gameId)) + 
  #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  theme(legend.position = "none")+
  labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId") + ylim(0,1)

ggsave(paste0(image_location, "/accuracy.png"))
```

-----

Everything here has bootstrapped 95% CIs.

Should find better curves to fit, but using quadratic to allow for some curvature.

```{r}
# ggplot(d.chat, aes(x=repNum, y=total_num_words, color=role))+
#   facet_wrap(~tangram, nrow=2)+
#   scale_color_brewer(palette="Dark2")+
#      stat_summary(fun.data = "mean_cl_boot")+
#   labs(title="Number of words", y="Number of words", x="Round number")+
#   theme(legend.position="bottom")
ggplot(d.chat, aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
  facet_wrap(~role, nrow=1)+
  scale_color_brewer(palette="Dark2")+
    geom_jitter(alpha=.05)+
    geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  #geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
     stat_summary(fun.data = "mean_cl_boot")+
    scale_y_continuous(limits = c(0,50))+
  labs(title="Number of words", y="Number of words", x="Round number", color="Player count")+
  theme(legend.position="bottom")
ggsave(here(image_location, 'words.pdf'), width=6, height=4)
```

```{r}
d.chat %>% filter(role=="speaker") %>% 
    mutate(groupxtangram=str_c(gameId,tangram)) %>% 
  group_by(repNum, numPlayers, gameId,tangram, groupxtangram) %>% 
  summarize(words=sum(total_num_words)) %>% 
ggplot(aes(x=repNum, y=words, color=as.factor(numPlayers)))+
  facet_wrap(~numPlayers, nrow=1)+
  scale_color_brewer(palette="Dark2")+
    geom_line(aes(group=groupxtangram), alpha=.1,method=glm, se=F)+
    geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
    #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
  labs(title="Words from speaker per tangram", y="Number of words", x="Round number", color="Player count")+
  theme(legend.position="null")
ggsave(here(image_location, 'words_lines.pdf'), width=6, height=4)
```

```{r}
d.chat %>% filter(role=="speaker") %>% 
ggplot(aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
  facet_wrap(~tangram)+
  scale_color_brewer(palette="Dark2")+
    #geom_smooth(method=glm, formula=y~poly(x,2), se=T, alpha=.1)+
      geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
       stat_summary(fun.data = "mean_cl_boot", size=.2)+
  labs(title="Tangram variability", y="Number of words", x="Round number", color="Player count")+
  theme(legend.position="bottom")
ggsave(here(image_location, 'words_tangrams.pdf'), width=8, height=6)
```

```{r accuracy}
d.round_results %>% group_by(playerId,repNum, gameId, numPlayers) %>% 
  mutate(correct.num=ifelse(correct,1,0)) %>% 
  ggplot(aes(x=repNum, y=correct.num, color=as.factor(numPlayers)))+
geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
  #geom_point()+
  scale_color_brewer(palette="Dark2")+
  #scale_y_continuous(limits = c(0,1))+
  labs(x="Round Number", y="Fraction correctly selected", title= "Overall accuracy increases over repetitions", color="Player count")+
    theme(legend.position="bottom")
ggsave(here(image_location, 'accuracy.pdf'), width=6, height=4)
```

```{r time}
d.round_results %>% group_by(playerId, repNum, gameId, numPlayers) %>% 
  filter(correct==T) %>% 
  #summarize(time=mean(time)) %>% 
  ggplot(aes(x=repNum, y=time, color=as.factor(numPlayers)))+
  geom_jitter(width=.4, height=0, alpha=.03)+
geom_smooth(method = "glm", formula = y~x,
                      method.args = list(family = gaussian(link = 'log')))+
  stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
  scale_y_continuous(limits = c(0,180))+
    scale_color_brewer(palette="Dark2")+
  labs(x="Round Number", y="Time to selection in seconds",
       title="People choose faster in later rounds", color="Player count")+
  theme(legend.position = "bottom")
ggsave(here(image_location, 'time.pdf'), width=6, height=4)
```
# Models

```{r, include=F}
model_input <- d.chat %>% filter(role=="speaker") %>% 
  mutate(block=repNum,
         words=total_num_words,
         tangram_group=str_c(tangram, gameId))

priors <- c(
  set_prior("normal(20, 20)", class="Intercept"),
  set_prior("normal(0, 10)", class="b"),
  set_prior("normal(0, 10)", class="sd"),
  set_prior("lkj(1)",       class="cor"))
model <- brm(words ~ block * numPlayers + (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "model_1"),                            prior=priors, control=list(adapt_delta=.95))

```

```{r, include=F}
d.prev.speaker <- d.chat %>% ungroup() %>%  filter(role=="speaker") %>% select(gameId,repNum, tangram, total_num_words_prev=total_num_words)
d.prev.round <- d.chat %>% ungroup() %>% select(playerId, correct, tangram, gameId, repNum) %>% 
  left_join(d.prev.speaker) %>% unique() %>% mutate(repNum=repNum+1)
d.chat.lagged <- d.chat %>%
  ungroup() %>% 
  select(gameId, playerId, trialNum, repNum, playerId, role, tangram, total_num_words, numPlayers) %>%
  left_join(d.prev.round) %>%
  mutate(reduction_word=log(total_num_words)-log(total_num_words_prev)) %>%
  filter(repNum>0) %>%
  filter(role=="speaker") %>%
  mutate(prev_correct_round=correct)
model_input <- d.chat.lagged %>% filter(role=="speaker") %>% 
  mutate(block=repNum,
         words=total_num_words,
         tangram_group=str_c(tangram, gameId),
         was_INcorrect=ifelse(!prev_correct_round,1,0))

priors <- c(
  set_prior("normal(20, 20)", class="Intercept"),
  set_prior("normal(0, 10)", class="b"),
  set_prior("normal(0, 10)", class="sd"),
  set_prior("lkj(1)",       class="cor"))
model_speaker_acc <- brm(words ~ block * numPlayers +block*was_INcorrect+ (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "speaker_acc"),                         prior=priors, control=list(adapt_delta=.95))

```

```{r}
summary(model)
```
```{r}
summary(model_speaker_acc)
```