### 2021-08-18 code/simulate_game.Rmd
move previous game data into data/version_0 and version preprocessing code

---
title: "simulate_game"
author: "Jessica Mankewitz"
date: "7/30/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r setup}
library(here)
source(here("code/theme_specs.R"))
library(png)
library(ggimage)
library(tidyverse)
```

```{r pressure, echo=FALSE}
d.players <- read_csv(here("data/processed_data/joined_data/players.csv"),
                      col_types = cols()) %>% distinct()
d.games <- read_csv(here("data/processed_data/joined_data/games.csv"),
                      col_types = cols()) %>% distinct()
d.rounds <- read_csv(here("data/processed_data/joined_data/rounds.csv"),
                      col_types = cols()) %>% distinct()
d.raw_chat <- read_csv(here("data/processed_data/joined_data/raw_chat.csv"),
                      col_types = cols()) %>% distinct()
d.contexts <- read_csv(here("data/processed_data/joined_data/contexts.csv"),
                      col_types = cols()) %>% distinct()
d.feedback <- read_csv(here("data/processed_data/joined_data/feedback.csv"),
                      col_types = cols()) %>% distinct()
d.demographics <- read_csv(here("data/processed_data/joined_data/demographics.csv"),
                      col_types = cols()) %>% distinct()

d.events <- read.csv(here("data/interpretable_transcript.csv"))
```

1) Figure out plotting t0 state of board (flowers available + maybe hidden markers?)
2) Figure out plotting t final state of board (selections)
3) Figure out plotting selections through events

For each round, need list of flowers available, assign coordinates?

```{r}
#available flowers are stored in d.contexts, which also has the image paths (woo!)
library(cowplot)

d.contexts %>%
  filter(gameId == "ExXQbCn35Xznf9tM5", trialNum == 0) %>%
  mutate(fullLocation = paste0(here(), "experiment/yellow_similar/image_06676.jpg")) %>%
  ggdraw() + draw_image(fullLocation)
```

#simulate utility structures?
# how often is each flower in the top 3 if 6 flowers are sampled from the 12 without replacement and the worth is assigned with a normal distribution?

```{r}
num_games <- 14400
#6 flowers for 24 trials for 100 games

utility_samples_low <- replicate(num_games,sample(1:17, size = 5, replace = FALSE, prob = 17:1))
dimnames(utility_samples_low) <- list(flower_num = 1:5, sample_num = 1:num_games)

flower_utility <- replicate(num_games,sample(18:20, size = 1, replace = FALSE))

d.utility_samples <- as.data.frame(as.table(utility_samples_low)) %>% 
  rename(flower_utility = Freq) %>% 
  mutate(flower_num = as.numeric(flower_num), 
         sample_num = as.numeric(sample_num)) %>%
  rbind(as.data.frame(flower_utility) %>% 
              mutate(flower_num = as.numeric(6), 
                     sample_num = as.numeric(1:num_games))) %>% arrange(sample_num, flower_num)

d.utility_samples$game <- rep(1:600, each=6*24)
```

```{r}
#flower samples

flower_samples <- replicate(num_games, sample(1:12, size = 6, replace = FALSE))
dimnames(flower_samples) <- list(flower_num = 1:6, sample_num = 1:num_games)

d.flower_samples <- as.data.frame(as.table(flower_samples)) %>% 
  rename(flower_name = Freq) %>% 
  mutate(flower_name = paste0("flower_", flower_name),
         flower_num = as.numeric(flower_num),
         sample_num = as.numeric(sample_num))

d.sampled_context = d.flower_samples %>% left_join(d.utility_samples) %>% rename(trial = sample_num)
```

#how often is each flower the highest?
```{r}

#top flower
d.sampled_context %>% group_by(game, trial) %>% 
  slice_max(order_by = flower_utility, n = 1) %>% 
  group_by(game, flower_name) %>%
  summarize(count = n()) %>%
  group_by(flower_name) %>%
  summarize(mean = mean(count),
            median = median(count))

#in the top 3
d.sampled_context %>% group_by(game, trial) %>% 
  slice_max(order_by = flower_utility, n = 3) %>% 
  group_by(game, flower_name) %>%
  summarize(count = n()) %>%
  group_by(flower_name) %>%
  summarize(mean = mean(count),
            sd = sd(count))

#estimated points (communication)
d.sampled_context %>% group_by(game, trial) %>% 
  slice_max(order_by = flower_utility, n = 3) %>% 
  summarize(trial_utility = sum(flower_utility)) %>%
  group_by(game) %>%
  summarize(mean_game_points = mean(trial_utility),
            sd_game_points = sd(trial_utility))

#penalty amount, if collision get n points
collision_penalty = 10
#estimated points with collisions
d.sampled_context %>% group_by(game, trial) %>%
  #take top 2
  slice_max(order_by = flower_utility, n = 2) %>% 
  #take smaller of top 2 (participant with the highest hidden)
  slice_min(order_by = flower_utility, n = 1) %>% 
  mutate(trial_utility = sum(flower_utility+(collision_penalty*2))) %>%
  group_by(game) %>%
  summarize(mean_game_points = mean(trial_utility),
            sd_game_points = sd(trial_utility))
```