### 2021-08-20 code/aa-flowers-pilot_analysis.Rmd
add nonlang conditions for comparison

---
title: "flowers-pilot-analysis"
output:
  html_document: default
  pdf_document: default
---

```{r set-up, message=FALSE, include=F}

library(jsonlite)
library(here)
library(tidyverse)

source(here("code/preprocessing/00_preprocessing_functions.R"))
source(here("code/theme_specs.R"))
#add pretty fonts
library(showtext)
font_add_google(name = "Assistant", family = "Assistant")
font_add_google(name = "Roboto", family = "Roboto")
showtext_auto()

##Data import constants
date_start=lubridate::ymd('2021-07-19')
fig_path=here("figs")
```

# Pre-process
Remove joined_data and uncomment the following block to rerun the preprocessing scripts
```{r process-data, message=FALSE, warning=FALSE}
data_list <- c("pilots", "collab_lang_20", "compet_lang_20", "nonlang_pilots", "compet_nonlang_15", "collab_nonlang_15")
process_all_data(data_list)
#join_all_data()
```

```{r read-processed-date, message=FALSE}
d.players <- read_csv(here("data/processed_data/joined_data/players.csv"),
                      col_types = cols()) %>% distinct()
d.games <- read_csv(here("data/processed_data/joined_data/games.csv"),
                      col_types = cols()) %>% distinct()
d.rounds <- read_csv(here("data/processed_data/joined_data/rounds.csv"),
                      col_types = cols()) %>% distinct()
d.raw_chat <- read_csv(here("data/processed_data/joined_data/raw_chat.csv"),
                      col_types = cols()) %>% distinct()
d.contexts <- read_csv(here("data/processed_data/joined_data/contexts.csv"),
                      col_types = cols()) %>% distinct()
d.feedback <- read_csv(here("data/processed_data/joined_data/feedback.csv"),
                      col_types = cols()) %>% distinct()
d.demographics <- read_csv(here("data/processed_data/joined_data/demographics.csv"),
                      col_types = cols()) %>% distinct()
```

Rename conditions for readability
```{r}
d.games <- d.games %>% mutate(conditionName = condition, 
                              condition = case_when(condition == "coopMulti" ~ "Cooperative",
                                         condition == "competCartel" ~ "Competitive",
                                         condition == "coopCartel" ~ "Old Cooperative"))
d.games
```

```{r}
#exclude BLOCKS with <12 completed (all 3 participants responded) rounds
# exclude incomplete games (not just incomplete rounds)

games_exclude <- d.rounds %>% group_by(gameId) %>% tally() %>% filter(n!=3*24) %>% select(gameId)

#exclude cooperative condition
d.rounds.final <-  d.rounds %>% left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
  filter(conditionName != "coopCartel") %>% anti_join(games_exclude)

d.raw_chat.final <- d.raw_chat %>% 
  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
  anti_join(games_exclude) %>% 
  write_csv(here("data/processed_data/joined_data/filtered_raw_chat.csv"))
```

Clean up a sharable data csv
```{r}
interpretable_data <- d.raw_chat.final %>% 
  add_chat_info(d.players, d.rounds.final, d.games) %>% 
    left_join(d.contexts %>% select(gameId, repNum, blockNum, trialNum, 
                                  playerResponse = label, 
                                  responseRawUtility = utility, blinded)) %>%
   #to avoid confusion, drop any round info thats actually a message
   mutate(playerResponse = ifelse(type == "message", NA, playerResponse),
         playerUtility = ifelse(type == "message", NA, playerUtility),
         responseRawUtility = ifelse(type == "message", NA, responseRawUtility),
         blinded = ifelse(type == "message", NA, blinded)) %>%
  select(gameId, trialNum, condition, chatEnabled, playerId, name, text, 
         participantAction, type, repNum, 
         blockNum, time, everything(), - submitted, -index, -createdAt) %>% write_csv(here("data/interpretable_transcript.csv"))
```

```{r}
#how many games are we left with
game_completion <- d.rounds %>% 
  group_by(gameId) %>% 
  summarize(num_rounds=max(trialNum+1)) %>% 
  left_join(d.games%>% ungroup() %>% select(gameId, condition, chat=chatEnabled, gameLength)) %>%
  mutate(gameComplete = ifelse(num_rounds == 24, TRUE, FALSE))

#knitr::kable(summary)
message("Full games")
game_completion %>% filter(num_rounds==24)
message("Partial games")
game_completion %>% filter(num_rounds!=24) 
```

Game counts
```{r}
#pull out pilots
d.rounds.final %>% left_join(d.games %>% select(gameId, chatEnabled)) %>% group_by(condition, chatEnabled) %>% 
  summarize(n_games = n_distinct(gameId)) %>% arrange(condition, chatEnabled)
```

# Demographics
```{r demographics}
d.demographics
```

```{r feedback}
d.feedback
d.feedback %>% ggplot(aes(x = human)) + geom_bar(stat = "count") + theme_jmank()
```

# Basic Analyses

Break out the by-player statistics for each round
```{r player-stats}

#player language per round
d.player_word_counts <- d.raw_chat.final %>% 
  filter(type == "message") %>%
  full_join(d.rounds.final, c("gameId", "trialNum", "repNum", 
                              "playerId", "numPlayers", "blockNum")) %>%
 # filter(!is.chitchat) %>% 
  mutate(text = gsub('[[:punct:] ]+',' ',text),
         text = str_squish(text),
         utt_length_chars = str_length(text), 
         utt_length_words = str_count(text, "\\W+") + 1) %>% 
  group_by(gameId, blockNum, trialNum, repNum, playerId, numPlayers) %>%
  summarize(text = paste0(text, collapse = ', '),
            total_num_words = sum(utt_length_words),
            total_num_chars = sum(utt_length_chars))

# player points per round
d.by_player_metrics <- d.rounds.final  %>% 
  left_join(d.player_word_counts) %>%
  left_join(game_completion %>% 
              select(gameId, gameComplete)) %>%
  left_join(d.games %>% select(-c(createdAt, flowers:scale), condition)) %>%  
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"))

```

Round info
```{r round-stats}
#get the # of raw points, max adjusted points, game legnth, # of words exchanged
d.max_round <- d.contexts %>%
  group_by(gameId, repNum, blockNum, trialNum) %>%
  #top 3 flowers
  slice_max(utility, n = 3) %>%
  summarise(roundMax = sum(utility))

d.by_round_metrics <- d.by_player_metrics  %>% 
  group_by(gameId, trialNum, repNum, blockNum) %>%
  summarize(groupPoints = sum(playerUtility, na.rm = T),
            groupNumWords = sum(total_num_words, na.rm = T),
            groupNumChars = sum(total_num_chars, na.rm = T),
            numFlowers = n_distinct(playerResponse)) %>%
  left_join(d.max_round) %>%
  left_join(game_completion %>% 
              select(gameId, gameComplete)) %>%
  left_join(d.games %>% select(-c(createdAt, flowers:scale), condition)) %>%  
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         roundPointsAdjusted = groupPoints/roundMax)
```

by block
```{r}
d.by_block_metrics <- d.by_round_metrics %>% 
  group_by(gameId, gameComplete, condition, 
           languageCondition, fullCondition, gameLength, blockNum) %>%
  summarize(blockPoints = sum(groupPoints, na.rm = T),
            blockNumWords = sum(groupNumWords, na.rm = T),
            blockNumChars = sum(groupNumChars, na.rm = T),
            blockMaxPoints = sum(roundMax),
            blockNumFlowers = sum(numFlowers)) %>%
  mutate(blockPointsAdjusted = blockPoints/blockMaxPoints,
         meanNumFlowers = blockNumFlowers*3/(6*3))
```

by game
```{r by-game}
d.by_game_metrics <- d.by_block_metrics %>%
  #from this point, only look at completed games
    filter(gameComplete) %>%
  #group by game
  group_by(gameId, condition, languageCondition, fullCondition, gameLength) %>%
  summarize(gamePoints = sum(blockPoints, na.rm = T),
            gameNumWords = sum(blockNumWords, na.rm = T),
            gameNumChars = sum(blockNumChars, na.rm = T),
            gameMaxPoints = sum(blockMaxPoints),
            gameNumFlowers = sum(blockNumFlowers)) %>%
  mutate(pointsAdjusted = gamePoints/gameMaxPoints,
         meanNumFlowers = gameNumFlowers*3/ (6*4*3))
```

```{r}
d.by_player_metrics %>% group_by(gameId, condition, languageCondition, playerId) %>% 
  summarize(mean_player_utility = mean(playerUtility)) %>%
  group_by(condition, languageCondition) %>% 
  summarize_at("mean_player_utility", funs(mean, min, max, sd))
```
## Simple Distributions
### Points

Unadjusted

```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>% 
  ggplot(aes(x = languageCondition, y = gamePoints, color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition)) + labs(x = "Language Condition", y = "Points Earned by Group", title = "Points Earned by Group") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")

ggsave(here(fig_path, "points_distirbution_raw.png"), width = 3, height = 5)
```

Max Adjusted

```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = pointsAdjusted, color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Language Condition", 
       y = "Points Earned by Group (Max Adjusted)", 
       title = "Points by Group") + 
  ylim(0,1) +
  scale_color_bay() + theme_jmank() + theme(legend.position = "none")

ggsave(here(fig_path, "points_distirbution_adjusted.png"),  width = 3, height = 5)
```

### Time
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = gameLength/60, 
             color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition))+ 
  labs(y = "Time (In Minutes)", title = "Game Length", x = "Condition")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "time_distirbution.png"),  width = 5, height = 3)

```

```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = log(gameLength/60), color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition)) + 
  ylim(0, 4.5)+
  labs(y = "Time (In Log Minutes)", title = "Game Length (Log Mins)")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")

ggsave(here(fig_path, "time_log_distirbution.png"))
```

### Number of Words
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = gameNumWords, color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition))+ 
  labs(y = "N. Words", title = "Number of Words Exchanged", x = "Condition")+scale_color_bay() + theme_jmank() +  theme(legend.position = "none")

ggsave(here(fig_path, "numwords_distirbution.png"), height = 5, width = 3)

```
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = languageCondition, y = log(gameNumWords+1), color = condition, shape = languageCondition)) + 
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  facet_grid(cols = vars(condition))+ 
  labs(y = "Log N. Words Exchanged")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "numwords_log_distirbution.png"))

```

## Covariate
### Time x Points
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(y = gamePoints, x = gameLength/60, color = condition, shape = languageCondition)) + 
  geom_point(alpha = .5) +
  geom_smooth(method=lm)+ 
  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+
  scale_color_bay() + theme_jmank() +  theme(legend.position = "none")
ggsave(here(fig_path, "time_words_scatter.png"))
```
# All Correlated
```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>%
  ggplot(aes(x = gameLength/60, y = gamePoints, shape = languageCondition, color = log(gameNumWords+1))) + 
  geom_point(alpha = .8) +
  geom_smooth(aes(group = condition), method=lm, color = "black")+
  facet_grid(cols = vars(condition)) + 
  labs(x = "Game length (in Minutes)", 
       y = "Points Earned by Group",
       color = "Log Num. Words",
       shape = "Language") + 
  theme_jmank() +  scale_color_steps(low = "#edd746", high = "#6B9555")

ggsave(here(fig_path, "all_correlated.png"), height = 3, width = 5)
```

```{r, fig.height=3, fig.width=5}
ms <- d.by_game_metrics %>%
  group_by(condition, languageCondition) %>%
  summarise(mean_len = mean(gameLength/60),
            ci_len = 1.96 * sd(gameLength/60) / sqrt(length(gameLength)),
            mean_points = mean(gamePoints),
            ci_points = 1.96 * sd(gamePoints) / sqrt(length(gameLength)))

d.by_game_metrics %>%
  ggplot(aes(x = gameLength/60, y = gamePoints, color = condition, shape = languageCondition)) +#, size = log(groupNumWords+1))) + 
  geom_smooth(aes(group = condition), method = "lm") +
  geom_point(alpha = .2) +
  geom_pointrange(data = ms, aes(x = mean_len, y = mean_points, 
                                 ymin = mean_points - ci_points,
                                 ymax = mean_points + ci_points)) +
  geom_errorbarh(data = ms, aes(x = mean_len, y = mean_points, 
                                 xmin = mean_len - ci_len,
                                 xmax = mean_len + ci_len), width = 0) + 
  # ylim(0,1500) + 
  xlim(0,90) + 
  labs(x = "Game length (in Minutes)", y = "Points Earned by Group")+
  ggthemes::scale_color_solarized() + 
  scale_color_bay() + theme_jmank() +  
  theme(legend.position = "bottom")

ggsave(here(fig_path, "time_points_scatter_nofacet.png"))

```

```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>% 
  ggplot(aes(x = log(gameLength/60), y = gamePoints, color = condition)) + 
  geom_point(alpha = .5) +
  geom_smooth(method=lm)+
  labs(x = "Game length (in Log Minutes)",y = "Points Earned by Group")+scale_color_bay() + theme_jmank()

ggsave(here(fig_path, "log_time_points_scatter_nofacet.png"))
```
### Language x Points

```{r}
ms <- d.by_game_metrics %>%
  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
  group_by(condition, fullCondition, languageCondition) %>%
  summarise(mean_nw = mean(gameNumWords),
            ci_nw = 1.96 * sd(gameNumWords) / sqrt(length(gameLength)),
            mean_points = mean(gamePoints),
            ci_points = 1.96 * sd(gamePoints) / sqrt(length(gameLength)))

d.by_game_metrics %>%
  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>% 
  #filter(langUse) %>%
  ggplot(aes(x = gameNumWords, y = gamePoints, color = condition, shape = languageCondition)) +
  geom_smooth(aes(group = fullCondition), method = "lm") +
  #geom_smooth(aes(group = fullCondition), method=glm, formula=y~poly(x,2), alpha=.3)+
  geom_point( alpha = .5) +
  geom_pointrange(data = ms, aes(x = mean_nw, y = mean_points,
                                ymin = mean_points - ci_points,
                                ymax = mean_points + ci_points)) +
  geom_errorbarh(data = ms, aes(x = mean_nw, y = mean_points,
                                xmin = mean_nw - ci_nw,
                                xmax = mean_nw + ci_nw), width = 0) +
  # ylim(0,1500) + 
  labs(x = "Words Exchanged", y = "Points Earned", title = ) + 
  scale_color_bay() + theme_jmank() +  
  theme(legend.position = "bottom")
```

```{r, fig.height=3, fig.width=5}
d.by_game_metrics %>% filter(languageCondition == "Lang") %>%
  mutate(langUse = ifelse(gameNumWords > 0, TRUE, FALSE)) %>%
  ggplot(aes(x= gameNumWords, y = pointsAdjusted, color = condition, shape = languageCondition)) + 
  geom_point(alpha = .5) +
  geom_smooth(method=lm)+ 
  labs(x = "N. Words Exchanged", y = "Group Points (Adjusted)")+scale_color_bay() + 
  theme_jmank() +  theme(legend.position = "none")

ggsave(here(fig_path, "nwords_points_scatter.png"), height = 3, width = 3)
```

### Progress Over Time

tried to make a gif? :/
```{r eval=FALSE, include=FALSE}
trial_num = 1
randomGame = sample(game_completion %>% filter(chat == T, gameComplete == T) %>% pull(gameId), 1)

save_plot <- function(rep_Num, block_Num, randomGame) {
  full_data <- d.by_round_metrics %>% 
    filter(repNum == rep_Num, blockNum == block_Num) %>%
  left_join(d.games %>% select(gameId, condition, chatEnabled)) %>% filter(chatEnabled)

  to_plot <- full_data %>% ggplot(aes(x = groupNumWords, y = roundPointsAdjusted, 
                                      shape = condition, color = condition)) + 
    geom_point(alpha = .2) + 
    geom_point(data = full_data %>% filter(gameId == randomGame), aes(x = groupNumWords, y = roundPointsAdjusted, 
                                      shape = condition, color = condition), alpha = 1) + 
    ggtitle(paste0("Block ", blockNum, ", Trial ", repNum))+
  theme_jmank() + scale_color_bay()

  ggsave(to_plot, file = here(fig_path, "animate", paste0(block_Num, "_", rep_Num, ".png")), height = 3, width = 5)
  }
#mapply(save_plot, seq(0,11), rep(0, 12), rep("sTX6ksbxqfnfsmzxe", 12))
```

Including language _optional_ games where they decided not to speak

```{r, fig.height=3, fig.width=5}
d.player_word_counts %>%
  left_join(d.games %>% 
              select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition, block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  ggplot(aes(x = repNum, y = groupNumWords, color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
ggsave(here(fig_path, "nwords_over_time.png"))
```

log (n words+1)
```{r, fig.height=3, fig.width=5}
d.player_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, condition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)+1) %>%
  ggplot(aes(x = repNum, y = log(groupNumWords), color = condition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + 
  labs(x = "Trial Number", y = "Log (N. Words + 1)")+ scale_color_bay() + theme_jmank() + theme(legend.position = "bottom")

ggsave(here(fig_path, "nwords_log_over_time.png"), height = 3, width = 5)
```

Only games where language was _used_

```{r, fig.height=3, fig.width=5}
d.player_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  filter(groupNumWords > 0) %>%
  ggplot(aes(x = repNum, y = groupNumWords, color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "N. Words")+ scale_color_bay() + theme_jmank()
```
log num words
```{r, fig.height=3, fig.width=5}
d.player_word_counts %>%left_join(d.games %>% 
                                   select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(languageCondition == "Lang") %>%
  group_by(gameId, fullCondition,block_name, trialNum, repNum) %>% 
  summarize(groupNumWords = sum(total_num_words, na.rm = T)) %>%
  filter(groupNumWords > 0) %>%
  ggplot(aes(x = repNum, y = log(groupNumWords+1), color = fullCondition)) + 
  geom_point(alpha = .2) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name)) + labs(x = "Trial Number", y = "Log N. Words")+ scale_color_bay() + theme_jmank()
```

### Accuracy
#### N. Flowers Selected each round
```{r}
d.by_round_metrics %>% 
  mutate(block_name = paste0("Block ", blockNum)) %>%
  ggplot(aes(x = repNum, y = numFlowers, color = condition, linetype = languageCondition)) + 
  geom_point(alpha = .1) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name), rows = vars(condition)) + labs(x = "Trial Number", y = "N. Unique Flowers")+ scale_color_bay() + theme_jmank()

ggsave(ggsave(here(fig_path, "collisions_conditionfacet.png")))
```
```{r}
d.by_round_metrics %>% 
  mutate(block_name = paste0("Block ", blockNum)) %>%
  ggplot(aes(x = repNum, y = numFlowers, color = condition, linetype = languageCondition)) + 
  geom_point(alpha = .1) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "N. Unique Flowers")+ scale_color_bay() + theme_jmank()

ggsave(ggsave(here(fig_path, "collisions_language_facet.png")))
```

Response Time
```{r}
d.rounds.final %>% group_by(gameId, repNum, blockNum, trialNum) %>% 
  summarize(mean_response_time = mean(time_sec, na.rm = T))  %>%
  left_join(d.games %>% select(gameId, condition,chatEnabled)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(condition != "coopCartel") %>%
  ggplot(aes(x = repNum, y = log(mean_response_time), color = condition, linetype = languageCondition)) + 
  geom_point(alpha = .1) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "log Mean response time") + scale_color_bay() + theme_jmank()
```

#More Accurate proportion possible/accuracy

Are they getting better over games?
```{r}
d.by_round_metrics %>%
  mutate(blockName = paste0("Block ", blockNum))%>% 
  ggplot(aes(x = groupNumWords, y = roundPointsAdjusted, 
             color = condition)) + 
  geom_point(alpha = .25) + theme_jmank() + scale_color_bay() + facet_wrap(~blockName) +
  labs(title = "Number of points per trial")
```
### Over/underpowered participants

```{r}
d.point_propotions <- d.by_player_metrics %>% 
  #points for entire game
  group_by(condition, languageCondition, gameId, playerId) %>% 
  summarize(player_total = sum(playerUtility)) %>% 
  group_by(condition, gameId) %>% mutate(game_total = sum(player_total)) %>% ungroup() %>%
  mutate(player_proportion = player_total/game_total)

#min
min_prop <- d.point_propotions %>% 
  group_by(condition, languageCondition, gameId) %>% 
  #lowest earned
  slice_min(order_by = player_proportion, n = 1) %>% 
  filter(condition == "Competitive") %>% select(-condition) %>% 
  mutate(player_placement = "Lowest Earning")

#max
max_prop <- d.point_propotions %>% 
  group_by(condition, languageCondition, gameId) %>% 
  slice_max(order_by = player_proportion, n = 1) %>% 
  filter(condition == "Competitive") %>% 
  select(-condition) %>% 
  mutate(player_placement = "Highest Earning")

rbind(min_prop, max_prop) %>%
  ggplot(aes(x = languageCondition, y = player_proportion, color = condition)) + 
  geom_hline(yintercept=.33, linetype = "dashed", alpha = .5)+
  geom_jitter(width = .1, alpha = .5) +
  facet_grid(cols = vars(player_placement)) + 
  stat_summary(fun.data = "mean_cl_boot")+
  labs(x = "Participant Placement", y = "Points earned by participant in place") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")

```
### Entropy

```{r}
entropy_func <- function(p1, p2, p3){
  return((p1 * log(p1)) + (p2 * log(p2)) + (p3 * log(p3)))
  #return((entropy_list[1] * log(entropy_list[1])) + (entropy_list[2] * log(entropy_list[2])) + (entropy_list[3] * log(entropy_list[3])))
}
d.game_entropies <- d.point_propotions %>% select(condition, languageCondition, gameId, player_proportion) %>% 
  mutate(p_num = rep(c("p1","p2","p3"), nrow(d.point_propotions)/3)) %>% 
  pivot_wider(names_from = p_num, values_from = player_proportion) %>% 
  mutate(entropy = 0-entropy_func(p1, p2, p3)) %>% select(-c(p1,p2,p3))

d.game_entropies %>% filter(condition == "Competitive") %>% 
  ggplot(aes(x = languageCondition, y = entropy)) + 
  geom_hline(yintercept=(0-entropy_func(1/3, 1/3, 1/3)), linetype = "dashed", alpha = .5)+
  #minimum is tricky? this is 12, 1, and 2 which is the least fair a game _could possible_ be, but isnt adjusted per game
  #geom_hline(yintercept=(0-entropy_func(12/15, 1/15, 2/15)), linetype = "dashed", alpha = .5)+
  geom_jitter(width = .1, alpha = .5) +
  stat_summary(fun.data = "mean_cl_boot")+
  labs(x = "Condition", y = "Entropy of Player Scores (by game)") + scale_color_bay() + theme_jmank() + theme(legend.position = "none")
```

### Selecting Hidden Flowers?

```{r}
d.rounds.final %>% left_join(d.games %>% 
                                   select(gameId, condition, chatEnabled)) %>%
  left_join(d.contexts %>% select(gameId, blockNum, trialNum, repNum, playerResponse = label, blinded)) %>%
  mutate(select_blinded = playerId == blinded) %>% 
  group_by(gameId, blockNum, trialNum, repNum, chatEnabled, condition) %>%
  summarize(numSelectHidden = sum(select_blinded)) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"),
         block_name = paste0("Block ", blockNum)) %>%
    filter(condition != "coopCartel") %>%
  ggplot(aes(x = repNum, y = numSelectHidden, color = condition, linetype = languageCondition)) + 
  geom_point(alpha = .1) +
  geom_smooth(method=lm)+
  facet_grid(cols = vars(block_name), rows = vars(languageCondition)) + labs(x = "Trial Number", y = "N. Select Hidden")+ scale_color_bay() + theme_jmank()
```

# Language Results

```{r}
# d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse, gameComplete)) %>%
#   mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
#          full_condition = paste0(condition, "-", chat)) %>%
#   ggplot(aes(x=time_sec, y = as.numeric(playerUtility), color = full_condition)) + 
#   facet_grid(rows = vars(condition), cols = vars(langUse)) +
#   geom_point(alpha = .25) + 
#   labs(x = "log time (in seconds)",
#        y = "individual utility",
#        title = "Response time in sec and utility", subtitle =  "each participant, each round")
# 
# d.round_results.final %>% left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>%
#   mutate(chat = ifelse(chatEnabled, "lang", "nonlang"),
#          full_condition = paste0(condition, "-", chat)) %>%
#   ggplot(aes(x=log(time_sec), y = as.numeric(playerUtility), color = full_condition)) + 
#   facet_grid(rows = vars(condition), cols = vars(langUse)) +
#   geom_point(alpha = .25) + 
#   labs(x = "log time (in seconds)",
#        y = "individual utility",
#        title = "Response time in sec (log) and utility", subtitle =  "each participant, each round")
```

Accuracy
sum utility for entire group/max possible utility per group (or, the max flower *3)
Each condition is going to have a different maximum utility

coopcartel will have a max utility = max flower * 3 (players) * 3 (incentive)

competcartel will have a max utility = sum(top_3_flowers)
```{r}
# d.max_utility <- d.contexts %>% 
#   left_join(d.games %>% select(gameId, condition)) %>%
#   mutate(utility= as.numeric(utility)) %>%
#   group_by(gameId, blockNum, repNum) %>% 
#   slice_max(order_by =utility,n = 3)%>%
#   summarize(competCartel = sum(utility),
#             coopMulti= sum(utility),
#             coopCartel = max(utility * 9)) %>% ungroup() %>%
#   pivot_longer(cols = c(competCartel, coopMulti, coopCartel), names_to = "condition", values_to = "maxUtility")
# 
# 
# d.utility <- d.round_results.final %>%
#          left_join(d.games %>% select(gameId, condition, chatEnabled, langUse)) %>% 
#   group_by(gameId, blockNum, repNum, trialNum,condition, langUse) %>% 
#   summarize(group_utility = sum(as.numeric(playerUtility)))%>% ungroup() %>%
#   left_join(d.max_utility) %>%
#   mutate(prop_utility = group_utility/maxUtility,
#          langUsed = if_else(langUse, "Lang", "Nonlang"))
# 
# d.utility %>% 
#   ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
#   geom_point(alpha = .4) + 
#   facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
#   #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
#   theme(legend.position = "none")+
#   labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
# 
# ggsave(paste0(image_location, "/accuracy.png"))
# 
# d.utility %>% 
#   ggplot(aes(x = trialNum, y=prop_utility, color = as.factor(gameId))) + 
#   geom_point(alpha = .4) + 
#   facet_grid(cols = vars(condition), rows = vars(langUsed), scales = "free_y") + 
#   #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
#   theme(legend.position = "none")+
#   labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId")
# 
# ggsave(paste0(image_location, "/accuracy.png"))
# 
# d.utility %>% 
#   filter(langUsed == "Lang") %>%
#   ggplot(aes(x = repNum, y=prop_utility, color = as.factor(gameId))) + 
#   geom_point(alpha = .4) + 
#   facet_grid(cols = vars(blockNum), rows = vars(gameId)) + 
#   #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
#   theme(legend.position = "none")+
#   labs(title="Accuracy", y="Group utility/max utility", x="Round number", color="gameId") + ylim(0,1)
# 
# ggsave(paste0(image_location, "/accuracy.png"))
```

-----

Everything here has bootstrapped 95% CIs.

Should find better curves to fit, but using quadratic to allow for some curvature.

```{r}
# ggplot(d.chat, aes(x=repNum, y=total_num_words, color=role))+
#   facet_wrap(~tangram, nrow=2)+
#   scale_color_brewer(palette="Dark2")+
#      stat_summary(fun.data = "mean_cl_boot")+
#   labs(title="Number of words", y="Number of words", x="Round number")+
#   theme(legend.position="bottom")
# ggplot(d.chat, aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
#   facet_wrap(~role, nrow=1)+
#   scale_color_brewer(palette="Dark2")+
#     geom_jitter(alpha=.05)+
#     geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
#   #geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
#      stat_summary(fun.data = "mean_cl_boot")+
#     scale_y_continuous(limits = c(0,50))+
#   labs(title="Number of words", y="Number of words", x="Round number", color="Player count")+
#   theme(legend.position="bottom")
# ggsave(here(image_location, 'words.pdf'), width=6, height=4)
```

```{r}
# d.chat %>% filter(role=="speaker") %>% 
#     mutate(groupxtangram=str_c(gameId,tangram)) %>% 
#   group_by(repNum, numPlayers, gameId,tangram, groupxtangram) %>% 
#   summarize(words=sum(total_num_words)) %>% 
# ggplot(aes(x=repNum, y=words, color=as.factor(numPlayers)))+
#   facet_wrap(~numPlayers, nrow=1)+
#   scale_color_brewer(palette="Dark2")+
#     geom_line(aes(group=groupxtangram), alpha=.1,method=glm, se=F)+
#     geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
#     #geom_smooth(method=glm, formula=y~poly(x,2), alpha=.3)+
#   labs(title="Words from speaker per tangram", y="Number of words", x="Round number", color="Player count")+
#   theme(legend.position="null")
# ggsave(here(image_location, 'words_lines.pdf'), width=6, height=4)
```

```{r}
# d.chat %>% filter(role=="speaker") %>% 
# ggplot(aes(x=repNum, y=total_num_words, color=as.factor(numPlayers)))+
#   facet_wrap(~tangram)+
#   scale_color_brewer(palette="Dark2")+
#     #geom_smooth(method=glm, formula=y~poly(x,2), se=T, alpha=.1)+
#       geom_smooth(method = "glm", formula = y~x,method.args = list(family = gaussian(link = 'log')))+
#        stat_summary(fun.data = "mean_cl_boot", size=.2)+
#   labs(title="Tangram variability", y="Number of words", x="Round number", color="Player count")+
#   theme(legend.position="bottom")
# ggsave(here(image_location, 'words_tangrams.pdf'), width=8, height=6)
```

```{r accuracy}
# d.round_results %>% group_by(playerId,repNum, gameId, numPlayers) %>% 
#   mutate(correct.num=ifelse(correct,1,0)) %>% 
#   ggplot(aes(x=repNum, y=correct.num, color=as.factor(numPlayers)))+
# geom_smooth(method = "glm", method.args = list(family = "binomial")) + 
#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
#   #geom_point()+
#   scale_color_brewer(palette="Dark2")+
#   #scale_y_continuous(limits = c(0,1))+
#   labs(x="Round Number", y="Fraction correctly selected", title= "Overall accuracy increases over repetitions", color="Player count")+
#     theme(legend.position="bottom")
# ggsave(here(image_location, 'accuracy.pdf'), width=6, height=4)
```

```{r time}
# d.round_results %>% group_by(playerId, repNum, gameId, numPlayers) %>% 
#   filter(correct==T) %>% 
#   #summarize(time=mean(time)) %>% 
#   ggplot(aes(x=repNum, y=time, color=as.factor(numPlayers)))+
#   geom_jitter(width=.4, height=0, alpha=.03)+
# geom_smooth(method = "glm", formula = y~x,
#                       method.args = list(family = gaussian(link = 'log')))+
#   stat_summary(fun.data = "mean_cl_boot", position = position_dodge(width=.2))+
#   scale_y_continuous(limits = c(0,180))+
#     scale_color_brewer(palette="Dark2")+
#   labs(x="Round Number", y="Time to selection in seconds",
#        title="People choose faster in later rounds", color="Player count")+
#   theme(legend.position = "bottom")
# ggsave(here(image_location, 'time.pdf'), width=6, height=4)
```
# Models

```{r, include=F}
# model_input <- d.chat %>% filter(role=="speaker") %>% 
#   mutate(block=repNum,
#          words=total_num_words,
#          tangram_group=str_c(tangram, gameId))
#          
# priors <- c(
#   set_prior("normal(20, 20)", class="Intercept"),
#   set_prior("normal(0, 10)", class="b"),
#   set_prior("normal(0, 10)", class="sd"),
#   set_prior("lkj(1)",       class="cor"))
# model <- brm(words ~ block * numPlayers + (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "model_1"),                            prior=priors, control=list(adapt_delta=.95))

```

```{r, include=F}
# d.prev.speaker <- d.chat %>% ungroup() %>%  filter(role=="speaker") %>% select(gameId,repNum, tangram, total_num_words_prev=total_num_words)
# d.prev.round <- d.chat %>% ungroup() %>% select(playerId, correct, tangram, gameId, repNum) %>% 
#   left_join(d.prev.speaker) %>% unique() %>% mutate(repNum=repNum+1)
# d.chat.lagged <- d.chat %>%
#   ungroup() %>% 
#   select(gameId, playerId, trialNum, repNum, playerId, role, tangram, total_num_words, numPlayers) %>%
#   left_join(d.prev.round) %>%
#   mutate(reduction_word=log(total_num_words)-log(total_num_words_prev)) %>%
#   filter(repNum>0) %>%
#   filter(role=="speaker") %>%
#   mutate(prev_correct_round=correct)
# model_input <- d.chat.lagged %>% filter(role=="speaker") %>% 
#   mutate(block=repNum,
#          words=total_num_words,
#          tangram_group=str_c(tangram, gameId),
#          was_INcorrect=ifelse(!prev_correct_round,1,0))
#          
# priors <- c(
#   set_prior("normal(20, 20)", class="Intercept"),
#   set_prior("normal(0, 10)", class="b"),
#   set_prior("normal(0, 10)", class="sd"),
#   set_prior("lkj(1)",       class="cor"))
# model_speaker_acc <- brm(words ~ block * numPlayers +block*was_INcorrect+ (block|tangram)+ (1|playerId)+(1|tangram_group)+(block|gameId), data=model_input,file=here(model_location, "speaker_acc"),                         prior=priors, control=list(adapt_delta=.95))

```

```{r}
#summary(model)
```
```{r}
#summary(model_speaker_acc)
```