### 2021-09-06 code/aa_mic_submission_analyses.Rmd
add prelim data tweaks and ms analyses

---
title: "aa_mic_submission_analyses"
author: "Jessica Mankewitz"
date: "9/6/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This notebook reads in the preprocessed and joined data, runs the basic processing/summarizing code, and generates figures and numbers for the 2021 MIC workshop submission.

```{r setup}
library(here)
library(tidyverse)

#for plotting
source(here("code/theme_specs.R"))

#Data import constants
date_start=lubridate::ymd('2021-07-19')
fig_path=here("figs")
read_data_path <- "data/processed_data/joined_data/"
```
# Summary Processing
## Read data

```{r read-data}
d.players <- read_csv(here(paste(read_data_path, "players.csv", sep = "")),
                      col_types = cols()) %>% distinct()
d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
                      col_types = cols()) %>% distinct()
d.rounds <- read_csv(here(paste(read_data_path, 
                                "rounds.csv", 
                                sep="")),
                      col_types = cols()) %>% distinct()
d.raw_chat <- read_csv(here(paste(read_data_path, 
                                  "raw_chat.csv", 
                                  sep="")),
                      col_types = cols()) %>% distinct()
d.contexts <- read_csv(here(paste(read_data_path, 
                                  "contexts.csv", 
                                  sep="")),
                      col_types = cols()) %>% distinct()
d.feedback <- read_csv(here(paste(read_data_path, 
                                  "feedback.csv", 
                                  sep = "")),
                      col_types = cols()) %>% distinct()
d.demographics <- read_csv(here(paste(read_data_path,
                                      "demographics.csv", 
                                      sep="")),
                      col_types = cols()) %>% distinct()
```

## Rename Conditions
Rename conditions for readability/alignment with ms
```{r}
d.games <- d.games %>% 
  mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName))
```

## Identify complete games

For the purposes of this ms, we're removing all unfinished games from the analyses 
```{r}
n_players = 3
n_rounds = 24
games_exclude <- d.rounds %>% group_by(gameId) %>% tally() %>% filter(n!=n_players*n_rounds) %>% select(gameId)

d.rounds.final <-  d.rounds %>% 
  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
  filter(conditionName != "coopCartel") %>% 
  mutate(game_complete = !(gameId %in% games_exclude$gameId))

d.raw_chat.final <- d.raw_chat %>% 
  left_join(d.games %>% select(gameId, conditionName, condition)) %>% 
  mutate(game_complete = !(gameId %in% games_exclude$gameId)) 
```

# Summary Stats
## Game counts
```{r}
d.games %>% 
  mutate(game_complete = !(gameId %in% games_exclude$gameId)) %>%
  group_by(-game_complete, condition, chatEnabled) %>%
  count()
```

## Game lengths
## Bonus Payments

# Points x Time x Num. Words

# Avg. Points Per Round

# Language Reduction

# Georgia Analyses