### 2022-06-10 code/prep_annotate.Rmd
process data for annotating flower referents

---
title: "prep_annotating"
author: "vboyce"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)

library(here)
library(tidyverse)

#for plotting

#Data import constants
date_start=lubridate::ymd('2021-07-19')
fig_path=here("figs")
read_data_path <- "data/processed_data/joined_data/"
```
# Summary Processing
## Read data

```{r read-data}
d.players <- read_csv(here(paste(read_data_path, "players.csv", sep = "")),
                      col_types = cols()) %>% distinct()
d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
                      col_types = cols()) %>% distinct()
d.rounds <- read_csv(here(paste(read_data_path, 
                                "rounds.csv", 
                                sep="")),
                      col_types = cols()) %>% distinct() 
d.raw_chat <- read_csv(here(paste(read_data_path, 
                                  "raw_chat.csv", 
                                  sep="")),
                      col_types = cols()) %>% distinct()
d.contexts <- read_csv(here(paste(read_data_path, 
                                  "contexts.csv", 
                                  sep="")),
                      col_types = cols()) %>% distinct()
d.feedback <- read_csv(here(paste(read_data_path, 
                                  "feedback.csv", 
                                  sep = "")),
                      col_types = cols()) %>% distinct()
d.demographics <- read_csv(here(paste(read_data_path,
                                      "demographics.csv", 
                                      sep="")),
                      col_types = cols()) %>% distinct()
```

## Rename Conditions
Rename conditions for readability/alignment with ms
```{r}
d.games <- d.games %>% 
  mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName))
```

## Identify complete games

For the purposes of this ms, we're removing all unfinished games from the analyses 
```{r}
n_players = 3
n_rounds = 24

game_completion <- d.rounds %>% 
  group_by(gameId) %>% 
  tally() %>% 
  mutate(game_complete = n==n_players*n_rounds) %>%
  rename(n_trials = n)

d.games.final <- d.games %>% 
  left_join(game_completion) %>% 
  replace_na(list(game_complete = FALSE, n_trials = 0))

d.games.conditions <- d.games.final %>% 
  select(gameId, color, condition, chatEnabled, game_complete) %>% 
  mutate(languageCondition = ifelse(chatEnabled, "Lang", "Nonlang"),
         fullCondition = paste(condition, languageCondition, sep = "-"))

d.rounds.final <-  d.rounds %>% 
  left_join(d.games.final %>% 
              select(gameId, conditionName, condition,color, chatEnabled, game_complete)) %>% 
  filter(conditionName != "coopCartel",
         game_complete)

d.raw_chat.final <- d.raw_chat %>% 
  left_join(d.games.final %>% 
              select(gameId, conditionName, condition, game_complete)) %>%
  filter(game_complete)

```

# Want to find a "best" game to start with

```{r}
d.rounds.final %>% filter(game_complete) %>% 
  group_by(gameId) %>% summarise(utils=sum(playerUtility)) %>% arrange(desc(utils))

# let's look at gameId == QHkfso2yMgbbziE7r first
```

```{r}

good_rounds <- d.rounds.final %>% filter(gameId=="QHkfso2yMgbbziE7r") %>% 
  select(playerId,playerResponse,roundId,gameId) %>% 
  separate(playerResponse, into=c("color","pResp"))

good_contexts <- d.contexts %>% filter(gameId=="QHkfso2yMgbbziE7r") %>% 
  select(roundId, gameId, label, utility) %>% 
  separate(label, into=c("color", "num")) %>% 
  mutate(labelutil=str_c(num,":",utility)) %>% 
  group_by(roundId, gameId, color) %>% 
  arrange(desc(utility)) %>% 
  summarize(flowers=str_c(labelutil,collapse=", "))

good_chat <- d.raw_chat.final %>% filter(gameId=="QHkfso2yMgbbziE7r") %>% 
  left_join(d.players %>% select(playerId, name)) %>% 
  rename(roundId=roundID) %>% 
  left_join(good_contexts) %>% 
  left_join(good_rounds) %>% 
  mutate(rowid=row_number()) %>% 
  select(rowid,trialNum,name, pResp, text, flowers, color) %>% 
  filter(!is.na(text)) %>% 
  write_tsv("test.tsv")

```