### 2022-09-22 code/annotation_analysis.Rmd
first pass at analysis

---
title: "Flowers text analysis"
author: "vboyce"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F, warning=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))

library(here)
library(tidyverse)

#for plotting
theme_set(theme_bw())
#Data import constants
date_start=lubridate::ymd('2021-07-19')
fig_path=here("figs")
read_data_path <- "data/processed_data/joined_data/"
annotation_path <- "data/annotations"
```

# VB to do notes
- Sbert
- compare with post-games stuff (and sbert that as well)
- consider being more consistent about what "the same one" or "that one" or etc got annotated
- inter-group differences/ correlations
- correlates of success

# Intro

I have annotated the transcripts of the full games marking what parts are referential (descriptive of a flower) and which flower (as best I can tell) it refers to.

See https://docs.google.com/presentation/d/1qcDRZzHbLhE-fp6W9nKcbGouj4w01qCLCy-yt072iiY/edit?usp=sharing for the flower number - image correspondences.

# Analysis

```{r}
d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
                      col_types = cols()) %>% distinct() |> 
    mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName)) |> 
  select(gameId,conditionName)

annotated <-  read_csv(here(annotation_path, "flowers_annotations.csv")) |> 
  filter(!is.na(span)) |> 
  mutate(numchar=str_length(span),
         numword=str_count(span, "\\W+") + 1) |> 
  left_join(d.games)
```

## How many words of referring language are there per round?

```{r}

sum_words <- annotated %>% group_by(trialNum, gameId,conditionName) %>% summarize(sum_char=sum(numchar),
                                           sum_word=sum(numword))

  ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
    #geom_line(aes(group=gameId))+
    facet_grid(.~conditionName)+
    geom_smooth(aes(group=gameId), method="lm", se=F)+
    geom_jitter(alpha=.5)

    ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
    geom_line(aes(group=gameId))+
    facet_grid(.~conditionName)

      ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
    #geom_line(aes(group=gameId))+
    facet_grid(.~conditionName)+
    geom_smooth(aes(group=gameId), method="lm", se=F)+
    geom_jitter(alpha=.5)

    ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
    geom_line(aes(group=gameId))+
    facet_grid(.~conditionName)

```
There are reduction trends (not sure if linear fit is correct or not) for most groups, some serious group differences, also noise.

## How long are individual referring expressions?

```{r}

  ggplot(annotated,aes(x=trialNum,y=numchar))+ geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)
  ggplot(annotated,aes(x=trialNum,y=numword))+geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)

```
Individual referring expressions tend to get shorter.

## How many referring expressions are there?
```{r}

annotated |> group_by(gameId, trialNum, conditionName) |> tally() |> 
  ggplot(aes(x=trialNum, y=n))+geom_jitter()+geom_smooth()+facet_grid(.~conditionName)

annotated |> group_by(gameId, referent, conditionName) |> tally() |> 
  group_by(conditionName) |> summarize(avg=mean(n), median=median(n))
```
Mostly the number of referring expressions is relatively constant, but the expressions themselves shorten.

# Sbert stuff
(currently copied from tangrams stuff)
```{r, include=F,}
sbert_concat <- read_csv(here(annotation_path,"pre_sbert_concat.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert_concat.RData')))  %>% as_tibble()

```

```{r, include=F,}
sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble()

```

```{r, include=F}

F_mat <- sbert_concat %>% select(starts_with("V")) %>% as.matrix() #Features
M_mat <- sbert_concat %>% select(-starts_with("V")) %>% mutate(feature_ind=row_number())
```

```{r helpers}
# note: cor expects features to be in columns so we transpose
get_sim_matrix = function(df, F_mat, method = 'cosine') {
  feats = F_mat[df$feature_ind,]
  if(method == 'cor') {
    return(cor(t(feats), method = 'pearson'))
  } else if (method == 'euclidean') {
    return(as.matrix(dist(feats, method = 'euclidean')))
  } else if (method == 'cosine') {
    return(as.matrix(lsa::cosine(t(feats))))
  } else {
    stop(paste0('unknown method', method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids) {
  ut <- upper.tri(cormat)
  data.frame(
    dim1 = ids[row(cormat)[ut]],
    dim2 = ids[col(cormat)[ut]],
    sim  = as.numeric(cormat[ut])
  ) %>%
    mutate(dim1 = as.character(dim1),
           dim2 = as.character(dim2))
}

make_within_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          .$repNum)) %>%
    mutate(rep1 = as.numeric(dim1), 
           rep2 = as.numeric(dim2)) 
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          as.character(.$combinedId)))
}
```