### 2022-10-14 code/annotation_analysis.Rmd
work on helper funcs for game x end sim (not entirely working!)

---
title: "Flowers text analysis"
author: "vboyce"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, message=F, warning=F)
knitr::opts_chunk$set(dev = "png", dev.args = list(type = "cairo-png"))

library(here)
library(tidyverse)
library(rstanarm)
library(rstan)
library(tidybayes)
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())
#for plotting
theme_set(theme_bw())
#Data import constants
date_start=lubridate::ymd('2021-07-19')
fig_path=here("figs")
read_data_path <- "data/processed_data/joined_data/"
annotation_path <- "data/annotations"
model_path <- "code/models"

show_summary_rstan <- function(model){
  badlist <- c("accept_stat__","divergent__","energy__","n_leapfrog__","sigma", "stepsize__","treedepth__")
  intervals <- gather_draws(model, `[a-zA-z_()]+`, regex=TRUE) |> mean_qi() |> filter(!.variable %in% badlist) |> 
    mutate(lower=round(.lower, digits=2),
           upper=round(.upper, digits=2),
           `Credible Interval`=str_c("[",lower,", ", upper,"]"),
           Term=.variable,
           Estimate=round(.value, digits=2)) %>% 
    select(Term, Estimate, `Credible Interval`)

  intervals
}

show_summary <- function(model){
  intervals <- gather_draws(model, `b_.*`, regex=T) %>% mean_qi()

  stats <- gather_draws(model, `b_.*`, regex=T) %>% 
    mutate(above_0=ifelse(.value>0, 1,0)) %>% 
    group_by(.variable) %>% 
    summarize(pct_above_0=mean(above_0)) %>% 
   mutate(`P-value equivalent` = signif(2*pmin(pct_above_0,1-pct_above_0), digits=2)) %>% 
    left_join(intervals, by=".variable") %>% 
    mutate(lower=round(.lower, digits=2),
           upper=round(.upper, digits=2),
           `Credible Interval`=str_c("[",lower,", ", upper,"]"),
           Term=str_sub(.variable, 3, -1),
           Estimate=round(.value, digits=2)) %>% 
    select(Term, Estimate, `Credible Interval`, `P-value equivalent`)

  stats
}

stats <- function(model, row){
  str_c(model[row,1],": Est=", model[row,2], ", CrI=", model[row,3])
}

stats_text  <- function(model, row){
  str_c( model[row,2], " (CrI=", model[row,3],")")
}
```

# Note

I have annotated the transcripts of the full games marking what parts are referential (descriptive of a flower) and which flower (as best I can tell) it refers to.

See https://docs.google.com/presentation/d/1qcDRZzHbLhE-fp6W9nKcbGouj4w01qCLCy-yt072iiY/edit?usp=sharing for the flower number - image correspondences.

# Cogsci notes

Key points:

- Many of the key reduction findings from tangrams generalize to this situation. Specifically, we see utterance reduction over time and w/i group convergence for each image and divergence between images. This situation is different in that we have different stimuli (more natural) and the set up is collaborative and more free-form in what is talked about. These patterns hold for both the individual and group payoff structures.

- One difference we see is that groups don't diverge (from each other). This may be dependent on stimulus properities (are there universal features of some of the images?) and group dynamics

- Conclusion: The key reference game findings have some generalizability. Settings like this one may be useful for encouraging discussion of a set of images and setting up partial knowledge situations.

# Analysis

```{r}
d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
                      col_types = cols()) %>% distinct() |> 
    mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName)) |> 
  select(gameId,conditionName)

annotated <-  read_csv(here(annotation_path, "flowers_annotations.csv")) |> 
  filter(!is.na(span)) |> 
  mutate(numchar=str_length(span),
         numword=str_count(span, "\\W+") + 1) |> 
  left_join(d.games) |> 
  mutate(flower=str_c(referent,color),
         player=str_c(gameId, name)) |> 
  select(gameId, player, flower, trialNum, pResp, normalized_span,color,numchar, numword, conditionName)
```

## How many words of referring language are there per round?

```{r}

sum_words <- annotated %>% group_by(trialNum, gameId,conditionName) %>% summarize(sum_char=sum(numchar),
                                           sum_word=sum(numword))

  # ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
  #   #geom_line(aes(group=gameId))+
  #   facet_grid(.~conditionName)+
  #   geom_smooth(aes(group=gameId), method="lm", se=F)+
  #   geom_jitter(alpha=.5)

    # ggplot(sum_words,aes(x=trialNum,y=sum_char, color=conditionName))+
    # geom_line(aes(group=gameId))+
    # facet_grid(.~conditionName)
    # 
    # 
      ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
    #geom_line(aes(group=gameId))+
    facet_grid(.~conditionName)+
    geom_smooth(aes(group=gameId), method="lm", se=F)+
    geom_jitter(alpha=.5)

    # ggplot(sum_words,aes(x=trialNum,y=sum_word, color=conditionName))+
    # geom_line(aes(group=gameId))+
    # facet_grid(.~conditionName)
    #

```
There are reduction trends (not sure if linear fit is correct or not) for most groups, some serious group differences, also noise.

## How long are individual referring expressions?

```{r}

 # ggplot(annotated,aes(x=trialNum,y=numchar))+ geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)
  ggplot(annotated,aes(x=trialNum,y=numword))+geom_jitter(alpha=.1)+ geom_smooth()+ facet_grid(.~conditionName)

```
Individual referring expressions tend to get shorter.

## How many referring expressions are there?
```{r}

annotated |> group_by(gameId, trialNum, conditionName) |> tally() |> 
  ggplot(aes(x=trialNum, y=n))+geom_jitter()+geom_smooth()+facet_grid(.~conditionName)
# 
# annotated |> group_by(gameId, referent, conditionName) |> tally() |> 
#   group_by(conditionName) |> summarize(avg=mean(n), median=median(n))
```
Mostly the number of referring expressions is relatively constant, but the expressions themselves shorten.

## Model

```{r,eval=F, include=F}

m1 <- stan_lmer(numword ~ trialNum + (1|gameId)+ (1|player)+(1|flower), data=annotated) |> write_rds(here(model_path,"m1.rds"))

```

```{r}
m1 <- read_rds(here(model_path,"m1.rds"))
m1
#prior_summary(m1)
show_summary_rstan(m1)
```

# Sbert stuff

Some pre-commentary on analytic choices:
- it's unclear whether we should be looking at referential statements individually (i.e. treating anything that occured as a separate utterance, after a line break, as separate) or concatenating everything Laju said about flower yellow3 into one statement. I think the first is generally better since sometimes a flower is described twice (perhaps in the same way), as in "the best values are 3 in a row and big fluffy" ... "okay, I'll take big fluffy". But there are also times when someone has a line break before continuing/clarifying a description. For now we go with single utterances.

- Unlike tangrams where there was a very clear cut every item gets described once every round by every group, here we have no such guarantees! So there's a big question of what scale to use for "time block". In some within game cases it might make sense to compare utterances within a single round (for distinctiveness of different descriptions or if multiple describe the same flower differently). It also might make sense to treat round as the semi-continuous variable that it is for some modelling where we can eventually see that farther away pairs do something. But for now, I'm going to divide the game into quarters (sets of 6 rounds) and pretend that we have 4 time points. This is arbitrary and preliminary but it's easy to do viz with.

```{r, include=F,}
sbert_concat <- read_csv(here(annotation_path,"pre_sbert_concat.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert_concat.RData')))  %>% as_tibble()

```

```{r, include=F,}
d.games <- read_csv(here(paste(read_data_path, "games.csv", sep = "")),
                      col_types = cols()) %>% distinct() |> 
    mutate(conditionName = condition, 
         condition = case_when(conditionName == "coopMulti" ~ "Shared Utilities",
                               conditionName == "competCartel" ~ "Individual Utilities",
                               conditionName == "coopCartel" ~ "Old Cooperative",
                               TRUE ~ conditionName)) |> 
  select(gameId,conditionName)

sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble() |> 
  left_join(d.games)

```

```{r, include=F}

F_mat <- sbert_single %>% select(starts_with("...")) %>% as.matrix() #Features
M_mat <- sbert_single %>% select(-starts_with("...")) %>% mutate(feature_ind=row_number()) |> 
  mutate(quarter=case_when(
    trialNum < 6 ~ 1,
    trialNum < 12 ~ 2,
    trialNum < 18 ~ 3,
    trialNum < 24 ~ 4
  ))
```

```{r helpers}
# note: cor expects features to be in columns so we transpose
get_sim_matrix = function(df, F_mat, method = 'cosine') {
  feats = F_mat[df$feature_ind,]
  if(method == 'cor') {
    return(cor(t(feats), method = 'pearson'))
  } else if (method == 'euclidean') {
    return(as.matrix(dist(feats, method = 'euclidean')))
  } else if (method == 'cosine') {
    return(as.matrix(lsa::cosine(t(feats))))
  } else {
    stop(paste0('unknown method', method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids) {
  ut <- upper.tri(cormat)
  data.frame(
    dim1 = ids[row(cormat)[ut]],
    dim2 = ids[col(cormat)[ut]],
    sim  = as.numeric(cormat[ut])
  ) %>%
    mutate(dim1 = as.character(dim1),
           dim2 = as.character(dim2))
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          as.character(.$combinedId)))
}
```

## Do flower labels diversify within games?

Hypothesis: descriptions for different flowers get more different over time (within game)

```{r }

flower_distinctive <- M_mat %>% 
  group_by(gameId,quarter, conditionName) %>% 
  mutate(combinedId=referent) %>% 
  make_across_df(F_mat, 'cosine') %>% 
  rename(referent1=dim1,referent2=dim2) %>% 
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  filter(referent1!=referent2) %>% 
  ungroup()
```

```{r}
ggplot(flower_distinctive, aes(x=quarter,y=sim,color=conditionName))+
      stat_summary(aes(group=str_c(gameId), color=conditionName),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.5, geom="point")+geom_smooth(formula=y~poly(x,2))
```

### TODO model
## Do groups diverge on labels?

Hypothesis: descriptions for the same flower between games get more different over time (maybe, for some flowers)

```{r }

game_distinctive <- M_mat %>% 
  group_by(quarter, color, referent) %>% 
  mutate(combinedId=gameId) %>% 
  make_across_df(F_mat, 'cosine') %>% 
  rename(gameId1=dim1,gameId2=dim2) %>% 
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  filter(gameId1!=gameId2) %>% 
  ungroup()
```

```{r}
ggplot(game_distinctive, aes(x=quarter,y=sim,color=color))+
      stat_summary(aes(group=str_c(color,referent), color=color),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.5, geom="point")+geom_smooth(formula=y~poly(x,2))
```

Wide variability -- some flowers might get more consistent names than others?
### TODO model

## Do group members converge with each other?

Hypothesis: players within the same game will label the same flower more similarly later in the game than earlier.

```{r }

player_similar <- M_mat %>% 
  group_by(quarter, color, referent, gameId) %>% 
  mutate(combinedId=name) %>% 
  make_across_df(F_mat, 'cosine') %>% 
  rename(name1=dim1,name2=dim2) %>% 
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  filter(name1!=name2) %>% 
  ungroup()
```

```{r}
ggplot(player_similar, aes(x=quarter,y=sim,color=color))+
      stat_summary(aes(group=str_c(color,referent), color=color),fun.data = "mean_cl_boot", position = position_dodge(width=.3), alpha=.5, geom="point")+geom_smooth(formula=y~poly(x,2))
```
### TODO model

# Within end-stuff

```{r}
sbert_post_test <- read_csv(here(annotation_path,"post_test.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert_post_test.RData')))  %>% as_tibble()

F_mat <- sbert_post_test %>% select(starts_with("...")) %>% as.matrix() #Features
M_mat <- sbert_post_test %>% select(-starts_with("...")) %>% mutate(feature_ind=row_number(), flower=str_c(color,num))
```

## Within flower
```{r}

end_distinctive <- M_mat %>% 
  group_by(flower, color) %>% 
  mutate(combinedId=str_c(gameId,"_",seen)) %>% 
  make_across_df(F_mat, 'cosine') %>% 
  separate(dim1, c("gameId1", "seen1")) |> 
  separate(dim2, c("gameId2", "seen2")) |> 
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  ungroup()
```

```{r}

end_distinctive |> mutate(same_group=ifelse(gameId1==gameId2, 
                                            "same_group","different_group"),
                          same_color=case_when(
  seen1=="own" & seen2=="own" ~ "own",
  seen2=="other" & seen1=="other" ~ "other",
  T ~"mixed"
))|> ggplot(aes(group=flower, x=color, y=sim, color=str_c(same_group,same_color)))+     
  stat_summary(aes(group=str_c(same_group, same_color)),fun.data = "mean_cl_boot", position = position_dodge(width=.3), geom="pointrange")

```
This is really poorly labelled. Needs stats. 
### TODO models

## Between flowers
```{r}

end_between <- M_mat %>% 
  group_by(color) %>% 
  mutate(combinedId=str_c(playerId, "_",gameId,"_",seen, "_", flower)) %>% 
  make_across_df(F_mat, 'cosine') %>% 
  separate(dim1, c("player1","gameId1", "seen1" ,"flower1")) |> 
  separate(dim2, c("player2", "gameId2", "seen2", "flower2")) |> 
  mutate(sim = ifelse(is.nan(sim), NA, sim)) %>%
  filter(flower1!=flower2) |> 
  ungroup()
```

```{r}

end_between |> mutate(same_group=case_when(
  player1==player2 ~ "same_person",
  gameId1==gameId2 ~ "same_group",
  T ~ "diferent_group"
), same_color=case_when(
  seen1=="own" & seen2=="own" ~ "own",
  seen2=="other" & seen1=="other" ~ "other",
  T ~"mixed"
))|> ggplot(aes(x=color, y=sim, color=str_c(same_group,same_color)))+     
  stat_summary(aes(group=str_c(same_group, same_color)),fun.data = "mean_cl_boot", position = position_dodge(width=.3), geom="pointrange")+facet_grid(.~same_group)

```
### TODO models

# TODO: between game & end-game

```{r}

d.raw_chat <- read_csv(here(paste(read_data_path, 
                                  "raw_chat.csv", 
                                  sep="")),
                      col_types = cols()) %>% distinct() |> 
  select(gameId, playerId) |> unique()

d.players <- read_csv(here(paste(read_data_path, "players.csv", sep = "")),
                      col_types = cols()) %>% distinct() |> select(playerId,name) |> inner_join(d.raw_chat)

sbert_single <- read_csv(here(annotation_path,"pre_sbert.csv")) %>% bind_cols(readRDS(here(annotation_path,'post_sbert.RData')))  %>% as_tibble() |> 
   left_join(d.players)
```

```{r to-end-helpers}

#here we only want sims between an in-game utt and an end description
# note: cor expects features to be in columns so we transpose

pairwiseCosine <- function(matrix1,matrix2) {
    apply(matrix1,2,function(x){
        apply(matrix2,2,lsa::cosine,x)
    })
}
get_sim_matrix = function(df, F_mat, method = 'cosine') {
  df_game <- df |> filter(type=="game")
  df_end <- df |> filter(type=="end")
  feats_game = F_mat[df_game$feature_ind,]
  print(t(feats_game))
  feats_end = F_mat[df_end$feature_ind,]
  print(t(feats_end))
  if(method == 'cor') {
    return(cor(t(feats_game), t(feats_end), method = 'pearson'))
  }  else if (method == 'cosine') {
    return(as.matrix(pairwiseCosine(t(feats_game), t(feats_end))))
  } else {
    stop(paste0('unknown method', method))
  }
}

# note this does de-duplicated version
flatten_sim_matrix <- function(cormat, ids_game, ids_end) {
  data.frame(
    dim1 = ids_end[row(cormat)],
    dim2 = ids_game[col(cormat)],
    sim  = as.numeric(cormat)
  ) %>%
    mutate(dim1 = as.character(dim1),
           dim2 = as.character(dim2))
}

make_across_df <- function(M_mat, F_mat, method) {
  M_mat %>%
    do(flatten_sim_matrix(get_sim_matrix(., F_mat, method = method),
                          as.character(filter(., type=="game")$combinedId), as.character(filter(., type=="end")$combinedId)))
}

```

** Does not work with grouped data frames, works with non-grouped dfs ** 
```{r}
#test
test_df1 <- tibble(color=c("blue","green"),id=c("game1","game2"), val1=c(.5,.7), val2=c(.9,.4)) |> mutate(type="game")
test_df2 <- tibble(color=c("blue", "green"),id=c("end1", "end2"), val1=c(1,-1), val2=c(-1,-1)) |> mutate(type="end")
test <- test_df1 |> union(test_df2) 
F_mat <- test %>% select(starts_with("val")) %>% as.matrix() #Features
M_mat <- test%>% select(-starts_with("val")) %>% mutate(feature_ind=row_number(), combinedId=id) |> group_by(color)

a <- make_across_df(M_mat, F_mat, method="cosine")

# want there to be one thing that gets grouped (M_mat) and then we separate it by type

sol <- tibble(dim1=c("end1","end2","end1","end2"), 
              dim2=c("game1", "game1", "game2","game2"),
              sim=c(lsa::cosine(c(.5,.9),c(1,-1)),
                    lsa::cosine(c(.5, .9), c(-1,-1)),
                    lsa::cosine(c(.7,.4),c(1,-1)),
                    lsa::cosine(c(.7,.4),c(-1,-1))))
```

# TODO: Correlates of success

# TODO: scale of variation between groups